Implementing Real-Time Transport Services over an Ossified Network

### 2016-05-23 paper.tex
subtitles to 2.; a few substantial edits; still working my way through it

%--------------------------------------------------------------------------------------------------
\subsection{Transport Services}

@@

Networked multimedia deadlines are unusual when compared to other real-time
systems. They are simultaneously flexible and strict: flexible in that the exact
value of the deadline is typically not important, provided it is of the right
order-of-magnitude for the application, but strict in that any particular
deadline provides a hard cut-off, after which the data is useless. \todo{this
para is largely unconvincing -- flexible is fine; strict is less believable, \ie
how are RT systems any different when the deadline expires? (Many RTSs will
discard, too.)}

The presence of deadlines, in a network that offers only a best-effort
packet delivery service, implies that the transport must provide
\emph{partial reliability}.
The network has non-zero probability of losing any particular packet.
If forward error correction is used to repair loss, this implies there is
always some probability that a particular packet will be non-recoverable.
If retransmission is used to recover from losses, there is the potential
for unbounded delay, since any retransmission may itself be lost.
These probabilities can be estimated, to give probabilistic constraints
on loss and delay, but will be non-zero. Accordingly, to meet deadlines,
the transport must support partial reliability service, accepting that
it may not be able to deliver all data by its deadline.
Many real-time applications run over TCP today, and TCP does not offer a
partial reliability service. This results in occasional play-out stalls,
when the transport blocks the application because a retransmission takes
too long. These stalls are one of the primary causes of poor user
experience in streaming applications, and are directly caused by the lack
of partial reliability in TCP. Missing one frame that has not been
delivered by its deadline is much less disruptive than a stall in play-out.

Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point in sending data
that cannot be used because it depends on previous data that was not
received. Providing this service is complicated by the two ways in which
data can be \emph{useful} to applications: it may itself by played out, or
it may be needed as part of the application's decoding chain. For example,
an I frame provides utility both in being played out, and in allowing for
subsequent P and B frames. As a result, dependency information may result
in deadline information being overruled: data may be sent even if it cannot
be played out, if it is necessary for the decoding of other data.

@@

We note that \emph{congestion control} is important. Clearly, the historic
belief that real-time applications require an isochronous channel, and
cannot be subjected to congestion control, is not practical on the
Internet. Further, while there is a number of applications that are not
adaptive (and require constant bandwidth), an increasing number are
adaptive, and would better serve users by adapting to the available
bandwidth.

>>>>

%--------------------------------------------------------------------------------------------------
\subsection{Timing and Transport Services}

@@

Networked multimedia deadlines are unusual when compared to other real-time
systems. They are simultaneously flexible and strict: flexible in that the exact
value of the deadline is typically not important, provided it is of the right
order-of-magnitude for the application, but strict in that any particular
deadline provides a hard cut-off, after which the data is useless. \todo{Is this para necessary at all? Its contribution is unclear and/or unconvincing -- flexible is fine; strict is less believable, \ie
how are RT systems any different when the deadline expires? (Many RTSs will
discard, too. Ultimately, I get what it is trying to say; I'm less sure it needs saying) Maybe the following(?):}
% When compared to conventional real-time systems, networked multimedia
% deadlines are neither soft nor hard.
It is useful to draw a distinction between in networked multimedia deadlines and
deadlines in conventional real-time systems. A multimedia deadline is neither soft, since arrivals played after a deadline contribute to poor experience; nor are deadlines hard, since occasional missed messages can be tolerated.

\subsection{Partial Reliability}
In a network that offers only a best-effort packet delivery service, deadlines
constrain delivery to \emph{partial reliability}.
%The network has non-zero probability of losing any particular packet.
For example, when used to repair loss, the limits of forward error correction
imply some probability that packet will be non-recoverable. By contrast,
retransmissions used to recover from loss have potentially unbounded delay
(since any retransmission may itself be lost).
% These probabilities can be estimated, to give probabilistic constraints on
% loss and delay, but will be non-zero. Accordingly, to meet deadlines the
Accordingly, a transport protocol that meets deadlines should provide partial
reliability, acknowledging that it may be unable to deliver all data by its
deadline.

Many real-time applications run over TCP today, though TCP offers no partial
reliability service. TCP's full reliability can lead to play-out stalls
when the application is blocked by retransmissions that take too
long. These stalls are one of the primary causes of poor user experience in
streaming applications.
% , and are directly caused by the lack of partial reliability in TCP.
For the applications under scrutiny, a missed frame that fails to deliver by its
deadline is much less disruptive than a stall in play-out.

\subsection{Message-oriented Dependencies}
Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point in sending data
that cannot be used because it depends on previous data that was not
received. Providing this service is complicated by the two ways in which
data can be \emph{useful} to applications: it may itself by played out, or
it may be needed as part of the application's decoding chain. For example,
an I frame provides utility both in being played out, and in allowing for
subsequent P and B frames. As a result, dependency information may result
in deadline information being overruled: data may be sent even if it cannot
be played out, if it is necessary for the decoding of other data.

@@

\subsection{Connection and Congestion Control}
We note that \emph{congestion control} is important. Clearly, the historic
belief that real-time applications require an isochronous channel, and
cannot be subjected to congestion control, is not practical on the
Internet. Further, while there is a number of applications that are not
adaptive (and require constant bandwidth), an increasing number are
adaptive, and would better serve users by adapting to the available
bandwidth.
